---
title: "Peter's data"
output: pdf_document
---

```{r setup, include=FALSE}
# Clear the workspace.
rm(list = ls())

# Load in data manipulation libraries.
library(tidyr)
library(dplyr)
library(lubridate)
library(ggplot2)

# Load in the raw serological cattle data. 
setwd("/Users/u1774615/GitHub/mosquito/data") # FIX ME
source("Peters_data_conversion_functions.R")
sero <- read.csv("raw/ndumo_cattle.csv", header=TRUE)

# Convert the date column to a date type.
sero$DATE <- as.Date(sero$date, format="%d%b%Y")

# Round down the animal age to:
# (i) an integer,
# (ii) maximum of age 9.
df <- sero %>%
    mutate(AGE = floor(age_y)) %>%
    mutate(AGE = ifelse(AGE > 9, 9, AGE))
```

## Data Summary

```{r summary, echo = FALSE}
# Summarize data
t <- data.frame(table(df$animal_id))
print(paste("number of samples:", length(df[,1])))
print(paste("number of unique animal IDs:", length(t[,1])))

# Some strange animal names, but looks unique:
# df[df$animal_id == "no tag 1s",] # follow up sequence looks good
# df[df$animal_id == "no tag 2s",] # follow up sequence looks good
# df[df$animal_id == "no tag 1",]  # rvf- and no follow-up
# df[df$animal_id == "no tag 2",]  # rvf- and no follow-up
# df[df$animal_id == "wrong tag",] # rvf- and no follow-up
# df[df$animal_id == "pinktag",] # rvf- and no follow-up
```

## Descriptive information about age

The median age at first capture is 4 years. This is largely consistent accross villages except the median ages at sites 1 and 6 are 5 years and 3 years respectively. Plot shows number of animals in each age.

```{r age, echo=FALSE}
# Age at first capture (gives only unique ids)
first <- df[df$DATE < as.Date("2016-09-09"), ]
# quantile(first$AGE) # median 4
# tapply(first$AGE, first$place, quantile) # largely consistent
ggplot(first, aes(AGE)) + geom_bar() + facet_wrap(~place)
```


## Age-prevalence and village-prev at first capture
Variation by village, but not a clear bias other than in sampling in village 3. 

```{r ageprev, echo=FALSE}
print(paste("overall prevalence at first capture:", 
length(first$rvf[first$rvf == 1])/length(first$rvf)))

#ggplot(first, aes(x= AGE)) + geom_bar() + facet_wrap(~rvf)
#tapply(first$AGE, first$rvf, quantile) # consistent

# Age-prev at first capture 
ggplot(first, aes(x = AGE, fill = as.factor(rvf))) + geom_bar()

# Prev at first capture by village
ggplot(first, aes(x = as.factor(place), fill = as.factor(rvf))) + geom_bar()

```

## Are capture periods distinct?
Yes, but for some sites more than others. Figure is counts by date, with second version only includes recapture periods. 
Note that villages 4 & 8 will only inform the model via initial captures.

```{r timing, echo=FALSE}
ggplot(df, aes(DATE)) + geom_histogram() + facet_wrap(~place) + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
ggplot(subset(df, df$DATE > as.Date("2016-11-20")), aes(DATE)) + 
    geom_histogram() + facet_wrap(~place) + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```


## Seroconversion
Number of new infections by diptank

```{r lump, echo=FALSE}
one <- get_one_village(1)
for(i in 2:9){
    t <- get_one_village(i)
    one <- rbind(one, t)
}
ggplot(one, aes(x = date.start, y = num_seroconvert, 
    fill = as.factor(diptankid))) + geom_col() + 
    theme(legend.position = "bottom")

ggplot(one, aes(x = date.start, y = num_tested, 
    fill = as.factor(diptankid))) + geom_col() + 
    theme(legend.position = "bottom")
```


## Guess capture periods that work for sites 5,6,7 9
Focus on these because 4 & 8 only include first capture
Sites 1 & 2 were rarely sampled so figure them out later!

```{r guess, echo=FALSE}
#t <- df[df$place %in% c(5, 6, 7, 9),]
t <- df[df$place %in% c(1, 2, 3, 5, 6, 7, 9),]

# unique(t$DATE); length(unique(t$DATE))
ggplot(subset(t, t$DATE > as.Date("2016-11-20")), 
       aes(DATE, fill = as.factor(place))) +
    geom_histogram(bins = 200) + #facet_wrap(~place) + 
    scale_x_date(date_breaks = "1 month") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r animal_id, echo=FALSE}
make_one_plot = function(place_number){
    t <- df[df$place %in% c(place_number), ]
    t$monthdate <- floor_date(t$DATE, 'month')
    if (place_number %in% c(1, 2, 3)){
        plotdf <- t
    } else{
        plotids <- t$animal_id[t$DATE > as.Date("2016-11-20")]
        plotdf <- subset(t, t$animal_id %in% plotids)
    }
    ggplot(plotdf,
       aes(x = monthdate, y = animal_id, fill = as.factor(rvf))) +
        geom_tile() + 
        scale_x_date(date_breaks = "1 month") + 
        theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), 
            axis.text.y = element_text(size = 6)) + 
        ggtitle(paste0("village: ", place_number))
}
make_one_plot(5)
make_one_plot(6)
make_one_plot(7)
make_one_plot(9)
```
